---
title: 'Statistical Inference Project 1: Part2 Basic Inferential Data Analysis'
author: "Frank Minjarez"
date: "3/13/2017"
output: html_document
keep_md: yes
---
Code Here: https://github.com/frankminjarez/StatisticalInference.git

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.


##  1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
library(datasets)
data(ToothGrowth)
head(ToothGrowth)
```

### The population of the dataset appears to have a normal distribution

```{r}
hist(ToothGrowth$len, 
     breaks = 15, 
     prob = TRUE, 
     main = "ToothGrowth Versus Normal Distribution", 
     xlab = "Observations")

## Sample density
lines(density(ToothGrowth$len), col="red")

## Theoretical normal density
xnorm <- seq(min(ToothGrowth$len), max(ToothGrowth$len), length = 15)
ynorm <- dnorm(xnorm, mean = mean(ToothGrowth$len), sd = sd(ToothGrowth$len))
lines(xnorm, ynorm, pch = 22, col = "blue", lty = 2)

legend("topright", c("ToothGrowth", "Normal"), 
       bty = "n", lty = c(1,2), col = c("red", "blue"))

str(ToothGrowth)
```
## 2. Provide a basic summary of the data
```{r}
library(ggplot2)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
table(ToothGrowth$supp, ToothGrowth$dose)
summary(ToothGrowth)

# Caclulate mean tooth len
meanToothLen <- aggregate(len~.,data=ToothGrowth,mean)
# Scatter plot and the avg lines for the average length
g <- ggplot(ToothGrowth,aes(x=dose,y=len))
g <- g + geom_point(aes(group=supp, colour=supp))
g <- g + scale_x_discrete("Dosage (Milligram)")
g <- g + scale_y_continuous("Length of Teeth") 
g <- g + geom_line(data=meanToothLen,aes(group=supp,colour=supp))
g + ggtitle("Scatter Plot: Effect of Dosage and Supplement on Tooth Growth")

g <- ggplot(ToothGrowth, aes(x=factor(dose),y=len, fill=factor(dose)))
g <- g + geom_boxplot(notch=F) + facet_grid(.~supp)
g <- g + scale_x_discrete("Dosage (Milligram)")
g <- g + scale_y_continuous("Length of Teeth")
g + ggtitle("Box Plot: Effect of Dosage and Supplement on Tooth Growth")
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

### Low p-value and higher mean len. OJ is more effective at 0.5 dosage.
```{r}

n <- subset(ToothGrowth, dose == 0.5)
t.test(len~supp, paired=F, var.equal=T, data=n)
```


### Low p-value and higher mean len. OJ is more effective at 1 dosage.
```{r}
n <- subset(ToothGrowth, dose == 1)
t.test(len~supp, paired=F, var.equal=T, data=n)
```

### High p-value and equal mean len. OJ and VC have the same effect at 2 dosage.

```{r}
n <- subset(ToothGrowth, dose == 2)
t.test(len~supp, paired=F, var.equal=T, data=n)
```

## 4. State your conclusions and the assumptions needed for your conclusions

Our early exploratory data analysis indicated the population has a normal distribution so we are assuming that we can use the t-interval test with our small sample set.

1. Both the scatter plot and the box plot indicate that as the dosage increases the effectiveness of the both the methods increases.
2. Hypothesis and confidence t-interval analysis indicate that for 0.5 and 1.0 dose values OJ is a more effective method. For dosage values of 2.0 both supplements achieve the same results.

```{r}
```
